Quantum Information

### 2024-12-05 examples/paper.tex
CP maps

@@

@@

@@

\begin{definition}[Entanglement]
  A state $\rho \in \bounded{\mathcal{H}} \otimes \bounded{\mathcal{K}}$ is called \emph{separable} if there are density matrices $\mu_i\in \bounded{\mathcal{H}}$, $\nu_i\in\bounded{\mathcal{H}}$, $i=1\dots n$, and a probability distribution $p\in\mathbb{R}^n$ such that
  \begin{equation*}
    \rho = \sum_{i=1}^n p_i \cdot \mu_i \otimes \nu_i.    
  \end{equation*}
  The state $\rho$ is called \emph{entangled} if it is not separable.
\end{definition}

\section{The maximally entangled state}

Let $\mathcal{H} = \mathbb{C}^d$ and consider the state $\ket{\Omega} = \frac{1}{\sqrt{d}}\sum_{i=0}^{d-1} \ket{ii}$. This state is called the maximally entangled state and it has the following properties.

>>>>

\subsubsection{The maximally entangled state}

Let $\mathcal{H} = \mathbb{C}^d$ and consider the state $\ket{\Omega} = \frac{1}{\sqrt{d}}\sum_{i=0}^{d-1} \ket{ii}\in\mathcal{H}\otimes\mathcal{H}$. This state is called the maximally entangled state and it has the following properties.

\begin{proposition}\label{prop:max_ent_partial_tr}
  \begin{equation*}
    (\id \otimes \tr )(\ket{\Omega}\bra{\Omega}) = (\tr \otimes\id)(\ket{\Omega}\bra{\Omega})  = \id.
  \end{equation*}
\end{proposition}

\begin{proof}
  Explicit calculation; exercise.
\end{proof}

@@

\begin{proof}
   $\rho$ is rank-one and Hermitian, thus it can be written as $\rho = \ket{\Psi}\bra{\Psi}$. $\tr \rho = \tr\ket{\Psi} \bra{\Psi} = \scalprod{\Psi}{\Psi} = \|\Psi\|^2 = 1$.
\end{proof}

@@

\begin{theorem}
   Let $\mathcal{H}$ and $\mathcal{K}$ be two Hilbert spaces, and $\ket{\Omega}\in\mathcal{H}\otimes\mathcal{H}$ be a maximally entangled state. Let $T:\bounded{\mathcal{H}}\to\bounded{\mathcal{K}}$ be linear. The following are equivalent:
   \begin{enumerate}
        \item $T$ is completely positive. \label{CP_Kraus_thm_1}
        \item $T$ is $\dim(H)$-positive. \label{CP_Kraus_thm_2}
        \item $(\id\otimes T)(\ket{\Omega}\bra{\Omega}) \geq 0$. \label{CP_Kraus_thm_3} 
        \item $T$ is Kraus. \label{CP_Kraus_thm_4}
   \end{enumerate}
\end{theorem}
\begin{proof}
  The implication \ref{CP_Kraus_thm_1} $\Rightarrow$ \ref{CP_Kraus_thm_2} $\Rightarrow$ \ref{CP_Kraus_thm_3} is by definition. To check \ref{CP_Kraus_thm_3} $\Rightarrow$ \ref{CP_Kraus_thm_4}, note that 
  \begin{equation*}
    T(\rho) = (\tr \otimes T) ((\id \otimes \rho ) \ket{\Omega}\bra{\Omega}) = (\tr \otimes T) ((\rho^T \otimes \id ) \ket{\Omega}\bra{\Omega}) = (\tr\otimes \id) ((\rho^T \otimes \id) X),
  \end{equation*}
  where $X = (\id\otimes T)(\ket{\Omega}\bra{\Omega})\geq 0$. As $X\geq 0$, we can write it as $\sum_i \ket{\Psi_i}\bra{\Psi_i}$ for some (unnormalized) vectors $\ket{\Psi_i}\in \mathcal{H}\otimes\mathcal{K}$. But $\ket{\Psi_i}$ then in turn can be written as $(\id \otimes A_i)\ket{\Omega}$, and thus $X$ can be written as 
  \begin{equation*}
    X = \sum_i (\id \otimes A_i) \ket{\Omega}\bra{\Omega} (\id \otimes A_i^\dagger). 
  \end{equation*} 
  Therefore 
  \begin{equation*}
    T(\rho) = (\tr\otimes \id) ((\rho^T \otimes \id) X) = \sum_i (\tr \otimes\id) ((\rho^T \otimes A_i) \ket{\Omega}\bra{\Omega} (\id\otimes A_i^\dagger)).
  \end{equation*}
  Using now \cref{prop:max_ent_pull_through} in the form  $(\rho^T \otimes \id)\ket{\Omega} = (\id\otimes \rho) \ket{\Omega}$, and \cref{prop:max_ent_partial_tr}, we obtain 
  \begin{equation*}
    T(\rho) = \sum_i (\tr \otimes\id) ((\id\otimes A_i \rho) \ket{\Omega}\bra{\Omega} (\id\otimes A_i^\dagger)) = \sum_i A_i \rho A_i^\dagger.
  \end{equation*}
  Finally to check \ref{CP_Kraus_thm_4} $\Rightarrow$ \ref{CP_Kraus_thm_1}, let the Kraus operators of $T$ be $A_i$, $n\in\mathbb{N}$ arbitrary, and $\rho\in\bounded{\mathcal{H}}\otimes \bounded{\mathbb{C}^n}$ be positive. Then
    \begin{equation*}
      (T\otimes \id_{\bounded{\mathbb{C}^n}})(\rho) = \sum_i (A_i \otimes \id_{\mathbb{C}^n}) \rho (A_i^\dagger \otimes \id_{\mathbb{C}^n} ) \geq 0,
    \end{equation*}
    i.e., $T$ is completely positive.  
\end{proof}

\begin{proposition}
  The transposition map is positive but not completely positive.
\end{proposition}
\begin{proof}
  Exercise.
\end{proof}

@@

\begin{definition}[Separability and entanglement]
  A state $\rho \in \bounded{\mathcal{H}} \otimes \bounded{\mathcal{K}}$ is called \emph{separable} if there are density matrices $\mu_i\in \bounded{\mathcal{H}}$, $\nu_i\in\bounded{\mathcal{H}}$, $i=1\dots n$, and a probability distribution $p\in\mathbb{R}^n$ such that
  \begin{equation*}
    \rho = \sum_{i=1}^n p_i \cdot \mu_i \otimes \nu_i.    
  \end{equation*}
  The state $\rho$ is called \emph{entangled} if it is not separable.
\end{definition}

\begin{proposition}
  Let $\mathcal{H}$ and $\mathcal{K}$ be Hilbert spaces, $\rho\in\bounded{\mathcal{H}}\otimes\bounded{\mathcal{K}}$ be a \emph{separable} state.  Let $T$ be a positive but not necessarily CP map. Then $(\id\otimes T)(\rho)\geq 0$.
\end{proposition}
\begin{proof}
  As $\rho$ is separable, there are density matrices $\mu_i\in \bounded{\mathcal{H}}$, $\nu_i\in\bounded{\mathcal{H}}$, $i=1\dots n$, and a probability distribution $p\in\mathbb{R}^n$ such that
    \begin{equation*}
      \rho = \sum_{i=1}^n p_i \cdot \mu_i \otimes \nu_i.    
    \end{equation*}
    Then 
    \begin{equation*}
      (\id\otimes T)(\rho) = \sum_{i=1}^n p_i \cdot \mu_i \otimes T(\nu_i) \geq 0,    
    \end{equation*}
    as $p_i$, $\mu_i$ and $T(\nu_i)$ are all positive.
\end{proof}

This can be used to detect entanglement: if $T$ is a positive but not necessarily CP map, and $\rho$ is a density matrix in a bipartite space, then if $(\id\otimes T)(\rho)\ngeq 0$, then we can conclude, apart from the fact that $T$ is not CP, that $\rho$ is entangled. A standard example is the transposition map.
